---
title: "SS3 Course ICES 2023"
author: "Mauricio Mardones"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
#bibliography: seaice.bib
#csl: apa.csl
link-citations: yes
linkcolor: blue
output:
  html_document:
    toc: true
    toc_deep: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: cosmo
    fontsize: 0.9em
    linestretch: 1.7
    html-math-method: katex
    self-contained: true
    code-tools: true
---

# Background

This code contains model files and associated material for the 2023 ICES course on Stock Synthesis held in Copenhagen, Denmark. 



# Model setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_knit$set(root.dir = "path/to/project/directory")
```

```{r message=FALSE}
# install.packages("devtools")
# devtools::install_github("r4ss/r4ss", ref="development")
# 
# install.packages("caTools")
# library("caTools")
# 
# install.packages("r4ss")
library(r4ss)
library(here)
#remotes::install_github("PIFSCstockassessments/ss3diags")
library(ss3diags)
library(knitr)
library(tidyverse)
```



```{r message=FALSE, warning=FALSE}
dir1<-here("herring-aspm")# version 3.30.21
dir2<-here("Workshops" , "SS3_ICES_Course", "herring-just-2019-as-equilibrium") 
dir3<-here("herring-scaa") # 
dir4<-here("Workshops" , "SS3_ICES_Course", "herring-scaa-2dar") # 
dir5<-here("herring-scaa-2dar-noTV") 
dir6<-here("Workshops" , "SS3_ICES_Course", "predators") # v. 3.30.21

```

Run models

```{r eval=FALSE, include=TRUE, warning=FALSE}
#shell(cmd="ss_win") # run SS para windows
# aqui debo cambiar el setwd() para correr el modelo
#dir5 <- setwd("~/DOCAS/SA_Krill/s5")
# r4ss::run(dirvec = dir3, model = './ss_osx', 
#                     skipfinished = FALSE) 
# #system('./ss_osx')

#another way to run with new r4ss

r4ss::run(
  dir = dir1,
  exe = "ss_osx",
  skipfinished = FALSE, # TRUE will skip running if Report.sso present
  show_in_console = TRUE # change to true to watch the output go past
)

r4ss::run(
  dir = dir1,
  exe = "ss_osx",
  skipfinished = FALSE, # TRUE will skip running if Report.sso present
  show_in_console = TRUE # change to true to watch the output go past
)

r4ss::run(
  dir = dir2,
  exe = "ss_osx",
  skipfinished = FALSE, # TRUE will skip running if Report.sso present
  show_in_console = TRUE # change to true to watch the output go past
)

r4ss::run(
  dir = dir3,
  exe = "ss_osx",
  skipfinished = FALSE, # TRUE will skip running if Report.sso present
  show_in_console = TRUE # change to true to watch the output go past
)

r4ss::run(
  dir = dir4,
  exe = "ss_osx",
  skipfinished = FALSE, # TRUE will skip running if Report.sso present
  show_in_console = TRUE # change to true to watch the output go past
)

r4ss::run(
  dir = dir5,
  exe = "ss_osx",
  skipfinished = FALSE, # TRUE will skip running if Report.sso present
  show_in_console = TRUE # change to true to watch the output go past
)

r4ss::run(
  dir = dir6,
  exe = "ss_osx",
  skipfinished = FALSE, # TRUE will skip running if Report.sso present
  show_in_console = TRUE # change to true to watch the output go past
)
```


# Diagnosis

## `herring-aspm` Model

```{r message=F, include=FALSE}
base.model1 <- SS_output(dir=dir1,
                        covar=F,
                        forecast=F)
```

Saco los outputs en html

```{r eval=FALSE,  message=F, include=FALSE}
SS_plots(base.model, 
         uncertainty = TRUE,
         datplot = F, 
         png=T, 
         aalresids = F,
         btarg=0.75, 
         minbthresh=0.25, 
         forecast=F)
```



Data disponible para este escenario. Espinel es la serie mas consistente del conjunto de datos.

```{r}
SSplotData(base.model1, subplot = 2)
```

Respecto a los valores y parametros biologicos modelados, los siguientes graficos identifican los estimadores puntuales del recurso

```{eval=FALSE}
SSplotBiology(base.model, 
              subplots =1, 
              labels = c("Length (cm)", "Age (yr)",
                         "Maturity", 
                         "Mean weight (kg) in last year",
                         "Spawning output",
                         "Length (cm, beginning of the year)", 
                         "Natural mortality",
                         "Female weight (kg)", 
                         "Female length (cm)", 
                         "Fecundity", 
                         "Default fecundity label",
                         "Year", 
                         "Hermaphroditism transition rate",
                         "Fraction females by age at equilibrium"),
 )
```

aporte de las cohortes por año para las capturas.
```{r}
SSplotCohortCatch(base.model1)
```

\quad

AJuste de tallas por flota
```{r}
SSplotComps(base.model1, subplots = 1)
```

Otros plots
```{r}
SSplotDynamicB0(base.model1, uncertainty = F)
#SSplotSPR(base.model3)
SSplotPars(base.model1)

```

Salida de las biomasas con las dos flotas


```{r}
SSplotTimeseries(base.model1, subplot = 1)
```


```{r}
SSplotTimeseries(base.model1,
  subplot=1,
  add = FALSE,
  forecastplot = FALSE,
  uncertainty = FALSE,
  bioscale = 1,
  minyr = -Inf,
  maxyr = Inf,
  plot = TRUE,
  print = FALSE,
  plotdir = "default",
  verbose = TRUE,
  btarg = "default",
  minbthresh = "default",
  xlab = "Year",
  labels = NULL,
  pwidth = 6.5,
  pheight = 5,
  punits = "in",
  res = 300,
  ptsize = 10,
  cex.main = 1,
  mainTitle = FALSE,
  mar = NULL
)
```

## `herring-scaa` Model

```{r message=F, include=FALSE}
base.model3 <- SS_output(dir=dir3,
                        covar=F,
                        forecast=F)
```

Saco los outputs en html

```{r eval=FALSE,  message=F, include=FALSE}
SS_plots(base.model3)
```



Data disponible para este escenario. Espinel es la serie mas consistente del conjunto de datos.

```{r}
SSplotData(base.model3, subplot = 2)
SSplotRecdevs(base.model3)
```

Respecto a los valores y parametros biologicos modelados, los siguientes graficos identifican los estimadores puntuales del recurso

```{r eval=FALSE}
SSplotBiology(base.model3, subplots =1, 
              labels = c("Length (cm)", 
                         "Age (yr)", 
                         "Maturity", 
                         "Mean weight (kg) in last year",
                         "Spawning output",
                         "Length (cm, beginning of the year)",
                         "Natural mortality",
                         "Female weight (kg)",
                         "Female length (cm)", 
                         "Fecundity", 
                         "Default fecundity label",
                         "Year", 
                         "Hermaphroditism transition rate", 
                         "Fraction females by age at equilibrium"),
               )
```

aporte de las cohortes por año para las capturas.
```{r}
SSplotCohortCatch(base.model3)
```

\quad

AJuste de tallas por flota
```{r}
SSplotComps(base.model3, subplots = 1)
```

Otros plots
```{r}
SSplotDynamicB0(base.model3, uncertainty = F)
#SSplotSPR(base.model3)
SSplotPars(base.model3)

```

Salida de las biomasas con las dos flotas


```{r}
SSplotTimeseries(base.model3, subplot = 10)
```


```{r eval=FALSE}
SSplotTimeseries(base.model3,
  subplot=1,
  add = FALSE,
  forecastplot = FALSE,
  uncertainty = FALSE,
  bioscale = 1,
  minyr = -Inf,
  maxyr = Inf,
  plot = TRUE,
  print = FALSE,
  plotdir = "default",
  verbose = TRUE,
  btarg = "default",
  minbthresh = "default",
  xlab = "Year",
  labels = NULL,
  pwidth = 6.5,
  pheight = 5,
  punits = "in",
  res = 300,
  ptsize = 10,
  cex.main = 1,
  mainTitle = FALSE,
  mar = NULL
)
```



## `herring-scaa-2dar-noTV` Model

```{r message=F, include=FALSE}
base.model5 <- SS_output(dir=dir5,
                        covar=F,
                        forecast=F)
```

Saco los outputs en html

```{r eval=FALSE,  message=F, include=FALSE}
SS_plots(base.model5, 
         uncertainty = TRUE,
         datplot = F, 
         png=T, 
         aalresids = F,
         btarg=0.75, 
         minbthresh=0.25, 
         forecast=F)
```



Data disponible para este escenario. Espinel es la serie mas consistente del conjunto de datos.

```{r}
SSplotData(base.model5, subplot = 2)
```

Respecto a los valores y parametros biologicos modelados, los siguientes graficos identifican los estimadores puntuales del recurso

```{r}
SSplotBiology(base.model5, subplots =1, 
              labels = c("Length (cm)", 
                         "Age (yr)", 
                         "Maturity", 
                         "Mean weight (kg) in last year",
                         "Spawning output",
                         "Length (cm, beginning of the year)",
                         "Natural mortality",
                         "Female weight (kg)",
                         "Female length (cm)", 
                         "Fecundity", 
                         "Default fecundity label",
                         "Year", 
                         "Hermaphroditism transition rate", 
                         "Fraction females by age at equilibrium"),
               )
```

aporte de las cohortes por año para las capturas.
```{r}
SSplotCohortCatch(base.model5)
```

\quad

AJuste de tallas por flota
```{r}
SSplotComps(base.model5, subplots = 1)
```

Otros plots
```{r}
SSplotDynamicB0(base.model5, uncertainty = F)
#SSplotSPR(base.model3)
SSplotPars(base.model5)

```

Salida de las biomasas con las dos flotas


```{r}
SSplotTimeseries(base.model5, subplot = 10)
```


```{r}
SSplotTimeseries(base.model5,
  subplot=1,
  add = FALSE,
  forecastplot = FALSE,
  uncertainty = FALSE,
  bioscale = 1,
  minyr = -Inf,
  maxyr = Inf,
  plot = TRUE,
  print = FALSE,
  plotdir = "default",
  verbose = TRUE,
  btarg = "default",
  minbthresh = "default",
  xlab = "Year",
  labels = NULL,
  pwidth = 6.5,
  pheight = 5,
  punits = "in",
  res = 300,
  ptsize = 10,
  cex.main = 1,
  mainTitle = FALSE,
  mar = NULL
)
```


## Desembarques


```{r include=FALSE}
start <- r4ss::SS_readstarter(file = file.path(dir1, "starter.SS"), 
                              verbose = FALSE)
dat <- r4ss::SS_readdat(file = file.path(dir1, start$datfile),
                        verbose = FALSE)
#ctl <- r4ss::SS_readctl(file = file.path(dir1, start$ctlfile),
#                        verbose = FALSE,
#                        use_datlist = TRUE, datlist = dat)

```

Los desembarques utilizados para cada una de las flotas que generan remoción en el recurso reineta, a saber; espinel, enmalle e industrial (Figura 6).

```{r echo=FALSE, fig.pos="H", fig.height=3,fig.width=7, fig.cap= "\\label{F2}Desembarques de merluza común por flota"}

cap <-dat$catch
cap1 <- ggplot(cap %>% 
                 filter(year!=-999), 
               aes(y=catch, x=year,
               fill=2)) +
  geom_bar(stat = 'identity', alpha=0.8) +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, hjust = 2))+
  #scale_fill_viridis(discrete=F)+
  #facet_wrap(.~fleet, labeller = as_labeller(flotas))+
  theme(legend.position="none")+
  ylab('Landings (t.)')+
  xlab('')+
  ggtitle('')
cap1
```

### Predator fleet with RW

Random Walk (RW) refers to a mathematical model that describes a stochastic process in which a variable changes randomly over time, without a clear trend or pattern.

Specifically, a random walk can be used as a Bayesian estimation technique to infer the posterior distribution of an unknown parameter. In this approach, it is assumed that the prior distribution of the parameter is a normal distribution with mean zero and a known variance, and that the parameter value at each time step follows a random walk process. Based on the observed data and the prior distribution, the posterior distribution of the parameter can be calculated using Bayesian inference.

Random walk is a useful tool for parameter estimation in dynamic models, as it allows for modeling uncertainty and variability in the parameter's evolution over time. However, it is important to note that the random walk assumes that the changes in the parameter are random and without a clear trend, which may not be appropriate in all cases.






```{r eval=FALSE}
#########################################################################
#COMPARACION DE MODELOS con distinto desembarque
#########################################################################


#PLOT labels, name of each model run
legend.labels <- c('herring-aspm',
                   'herring-just-2019-as-equilibrium')

#read in all model runs
#note if cover=T you need a hessian; if covar=F you do not need a hessian
biglist <- SSgetoutput(keyvec = NULL, 
                       dirvec = c(dir1,dir2,dir3),	
                       getcovar = F)

#create summary of model runs from list above
summaryoutput <- SSsummarize(biglist)

SSplotComparisons(summaryoutput, subplots = 1:20, plot = TRUE,
                  print = T, endyrvec = "default", indexfleets = NULL,
                  indexUncertainty = FALSE,
                  indexQlabel = TRUE, 
                  indexQdigits = 4, 
                  indexSEvec = "default",
                  indexPlotEach = FALSE, 
                  labels = c("Year", "Spawning biomass (t)",
                             "Relative spawning biomass", 
                             "Age-0 recruits (1,000s)",
                             "Recruitment deviations", 
                             "Index", 
                             "Log index", 
                             "1 - SPR", 
                             "Density",
                             "Management target", 
                             "Minimum stock size threshold",
                             "Spawning output",
                             "Harvest rate"), col = NULL, 
                  shadecol = NULL, pch = NULL,
                  lty = 1, lwd = 2, spacepoints = 10, staggerpoints = 1,
                  initpoint = 0, tickEndYr = TRUE, shadeForecast = TRUE,
                  xlim = "default", ylimAdj = 1, xaxs = "r", yaxs = "r",
                  type = "o", uncertainty = TRUE, 
                  shadealpha = 0.1, legend = TRUE,
                  legendlabels = "default", legendloc = "topright",
                  legendorder = "default", legendncol = 1, sprtarg = NULL,
                  btarg = NULL, minbthresh = NULL, 
                  pwidth = 6.5, pheight = 5,
                  punits = "in", res = 300, ptsize = 10,
                  plotdir = "~/DOCAS/Workshops/SS3_ICES_Course", 
                  filenameprefix = "", 
                  densitynames = c("SSB_Virgin", "R0"),
                  densityxlabs = "default", densityscalex = 1,
                  densityscaley = 1, densityadjust = 1, 
                  densitysymbols = TRUE,
                  densitytails = TRUE, 
                  densitymiddle = FALSE, 
                  densitylwd = 1,
                  fix0 = TRUE, new = TRUE, 
                  add = FALSE, 
                  par = list(mar = c(5, 4, 1, 1) + 0.1), 
                  verbose = TRUE, mcmcVec = FALSE,
                  show_equilibrium = TRUE)



SStableComparisons(summaryoutput,
                   likenames = c("TOTAL", 
                                 "Survey", 
                                 "Length_comp", 
                                 "Age_comp", 
                                 "priors",
                                 "Size_at_age"),
                   names = c("Recr_Virgin", 
                             "R0", "steep", 
                             "NatM",
                             "L_at_Amax", 
                             "VonBert_K", 
                             "SSB_Virg", 
                             "Bratio_2017", "SPRratio_2016"),
                   digits = NULL, modelnames = "default", csv = FALSE,
                   csvdir ="~/DOCAS/Workshops/SS3_ICES_Course", 
                   csvfile = "parameter_comparison_table.csv", verbose = TRUE,
                   mcmc = FALSE)
```

# Analisis retrospectivo

```{r eval=FALSE}

#do retrospective model runs
retro(dir=file.path('~/DOCAS/Workshops/SS3_ICES_Course',, olsubdir="",
      newsubdir = "retrospectives",
           subdirstart = "retro", years = 0:-5,
           overwrite = TRUE, exefile = "ss",
           extras = "-nox -nohess", 
           intern = FALSE, CallType = "system",
           RemoveBlocks = FALSE)

retroModels <- SSgetoutput(dirvec=file.path('~/DOCAS/Workshops/SS3_ICES_Course',
                                            "retrospectives",
                                            paste("retro",0:-5,sep="")))

retroSummary <- SSsummarize(retroModels)
endyrvec <- retroSummary$endyrs + 0:-5

SSplotComparisons(retroSummary, 
                  endyrvec=endyrvec, 
                  legendlabels=paste("Data",0:-5,"years"),
                  plotdir='~/DOCAS/Workshops/SS3_ICES_Course',
                  plot=TRUE,print=T)

TableCompare <- SStableComparisons(retroSummary,
                                   likenames=like, 
                                   names=names, 
                                   modelnames=c('B','-1','-2','-3','-4','-5'), 
                                   csv=TRUE, 
                                   csvfile="RetroRuns.csv",verbose=FALSE)

```



# Course Notes

- Lorenzen mortality works just in case you have parametric growth. Used a lot in US in Europe (Cardinale, 2023).
About F:
- Both use the same F calculations
- Hybrid best when F is low
- At high F, hybrid’s algorithm to maintain match to catch inhibits movement towards convergence
- Parameter approach can be configured to get starting value for F parameter during early phases using hybrid, then converts to parameter to finish the run
- F method = 4 gives best flexibility because it allows hybrid vs parameter conversion on a fleet-specific basis (today is the best method. If you have fleet specific).
- 0.23 # F ballpark value in units of annual_F. Is usefuul when we have outlier data (High biomas or high F) and you can set what year happen this, i. e. ; -2008 # F ballpark year (neg value to disable).
- Body weigth is in kilograms
-Options 3 & 4 are experimental in SS3, but match some other models, like small pelagics stock.
- If fixing without having an intended rationale, then let the estimated recruitments deviate from the curve:
–- use recdev option 2, not 1, for more flexibility
–- set lambda on recdevs to low value (<0.10)
- methods to data weighting  `tune_comps()`
- Abundance predator is lika a Fishing effort (F)

- run model with -nohess and 0 phasses

- time varing GP process is usefull when we have bad fits in some part of our

- About BRP
file `summary_sso` have a summary of outputs  like variables and Mgmt_Quantity

- nll colud be near 1000. Many millions is a wrong!



# Q&A

- I have length comosition by area, but nothing more, could use this info like fleets?
- I have size composition from  diets of predators. 

